Resolvable Ambiguity

### 2019-07-11 paper/paper.tex
Pass over Conclusion, adding future work

\section{Conclusion}
In this paper, we introduce the concept of \emph{resolvable ambiguity}. This means that a language syntax definition can be ambiguous, but the ambiguity is resolvable later by the end user when compiling a program. This approach departs from the common standpoint that grammars and syntax definitions of languages must be unambiguous. We motivate our work both for general-purpose languages in general, but for domain-specific languages in particular. As part of the new concept, we formalize the fundamental \emph{resolvable ambiguity problem}, and divide it into static and dynamic parts. We prove soundness and completeness of static resolvablity for a restricted class of languages, and soundness and completeness of dynamic resolvability for general context-free languages with balanced parenthesis. The approach is evaluated on two case studies, including the definition of a large subset of OCaml, and on the composition of domain-specific languages.

>>>>

\section{Conclusion}
In this paper, we have introduced the concept of \emph{resolvable
  ambiguity}. A language grammar is resolvably ambiguous if all
ambiguities can be resolved by the end-user at parse time. This
approach departs from the common standpoint that grammars and
syntax definitions of languages must be unambiguous.
%
As part of the new concept, we have formalized the fundamental
\emph{resolvable ambiguity problem}, divided it into static and
dynamic parts, and provided solutions for both variants for a
restricted class of languages. Through case studies, we have
showed practical applicability of the approach, both for building
new domain-specific languages and for reasoning about existing
general-purpose languages.

In future work, we will investigate weakning the restrictions
currently needed for the static and dynamic resolvability
algorithms to work. We also intend to develop our theory and tools
to support handling deep precedence ambiguities and ambiguities
based on ``longest match''. A long term goal is to be able to
suggest changes to the grammar of a language, based on ambiguities
found through dynamic analysis.
